Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-05-16 doc/GMP2015/revision_2/s06_problempinning.tex
Revision 2. 24 hours to submission.

\begin{figure}\centering
    \includegraphics[height=\myMinHeight]{img/Silica}
    \hspace{0.5cm}
    \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
    % \raisebox{0.25\height}{
    %     \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
    % }
    \caption{Example of a single monolayer of a silica (silicon and oxygen) glassy structure. This can be thought of as a triangle-multipin qusecs where the silicon atoms are the triangles and the oxygen atoms are the pins. Not shown here are other monolayers stacked on this one. Each silicon atom has the structure seen to the right, and so binds to another oxygen atom in an adjacent monolayer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
    \label{fig:silica_glass}
\end{figure}

We can use qusecs DR-plans to design materials such as disordered graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}. We will investigate a more specific problem in a somewhat more general setting: the problem of finding boundary conditions (additional constraints) to add to an underconstrained monolayer to make it isostatic. This can be done in a number of ways: (1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}); (2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or (3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).

In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress and other property design. To answer these questions, we first introduce the qusecs that are used to model these materials. We will discuss item (2) in detail in this section.

@@

\begin{definition}
    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies, some subsets of which are pinned together by pins.
\end{definition}

\begin{remark*}\label{rem:bodypin_is_barjoint}
    A body-hyperpin qusecs is a special case of bar-joint qusecs of the first 5 sections. As such, the DR-planning for isostatics discussed in Section \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
\end{remark*}

For the rest of this section, we will only be dealing with the DR-plan of such qusecs. As such, we will refer only to the underlying hypergraph of the qusecs. We now introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}, for which the optimal completion problem is significantly easier.

@@

\begin{definition}
    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares at most 3 pins with other bodies. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

@@

We can develop the notion of a canonical DR-plan when the hypergraph is $(k,l)$-sparse (i.e., combinatorially independent) using certain notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}, provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}). In particular, we can define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.

\begin{observation*}
\label{obs:bodypin_drp}
    For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
\end{observation*}

@@

The strong Church-Rosser property holds, making all canonical DR-plans optimal:
%
\begin{remark}\label{rem:1dofcanon}
    When the input is independent, all $(3,4)$-tight canonical DR-plans are optimal. We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}. This is because such $(3,4)$-tight graphs are matroidal and have a pebble game \cite{Lee:2007:PGA}.
\end{remark}

@@

\begin{observation*}
\label{obs:2dof_case}
    For the 2-dof case, if a similar statement to Remark \ref{rem:1dofcanon} can be made, then Theorem \ref{thm:1dofcase} is true for 2-dof as well.
\end{observation*}

@@

\begin{observation*}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer permit the algorithm given in the proof. In particular the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if that case were matroidal and an optimal DR-plan is known).
\end{observation*}

>>>>

\begin{figure}\centering
    \includegraphics[height=\myMinHeight]{img/Silica}
    \hspace{0.5cm}
    \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
    % \raisebox{0.25\height}{
    %     \includegraphics[height=\myMinHeight]{img/Silicon_tetrahedron}
    % }
    \caption{Example of a single monolayer of a silica (silicon and oxygen) glassy structure. This can be viewed as a triangle-multipin qusecs where the silicon atoms are the triangles and the oxygen atoms are the pins. Not shown here are other monolayers stacked on this one. Each silicon atom has the structure seen to the right, and so binds to another oxygen atom in an adjacent monolayer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
    \label{fig:silica_glass}
\end{figure}

We can use qusecs DR-plans to design materials such as disordered graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}. We investigate a more specific problem in a somewhat more general setting: the problem of finding boundary conditions (additional constraints) to add to an underconstrained monolayer to make it isostatic. This can be done in a number of ways: (1) pin together 2 underconstrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}); (2) pin the boundary of (or in general, add constraints to) a layer (possibly a genus 0 monolayer) so that it becomes isostatic; or (3) design a broader class of structures to ensure they are isostatic, self-similar (via some subdivision rule) and in addition isostatic at each level of the subdivision (see Figure \ref{fig:subdivision}).

In all cases, we are specifically interested in how to add additional constraints such that the resulting isostatic structure has a small DR-plan; this way a realization can be found, allowing efficient stress, flex and other property design related to the rigidity matrix. To answer these questions, we first introduce the qusecs' that are used to model these materials. In this section, we discuss Item (2) in detail.

@@

\begin{definition}
    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies,  subsets of which are pinned together by pins; i.e, are incident at a common point.
\end{definition}

\begin{remark*}\label{rem:bodypin_is_barjoint}
    A body-hyperpin qusecs is a special case of bar-joint qusecs of the previous sections of the paper. As such, the DR-planning for isostatic systems discussed in Section \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} still go through with minor modifications.
\end{remark*}

For the remainder of this section, we deal only with the DR-plan of such qusecs. Hence, we refer only to the combinatorics or underlying hypergraph of the qusecs. We now introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}, for which the optimal completion problem is significantly easier.

@@

\begin{definition}
    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body is a triangle, i.e., it shares  pins with at most 3 other bodies. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hypergraphs, {\em 1 and 2-dof} can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

@@

Provided such $(k,l)$-sparse graphs are matroidal (conditions given in \cite{Lee:2007:PGA}),
the notion of a canonical DR-plan extends directly to the case when the hypergraph is $(k,l)$-sparse (i.e., independent) using the straightforward notion of trivial and non-trivial intersections and $(k,l)$-tightness conditions as in Section \ref{sec:DRP}. In particular, we define canonical DR-plans with similar properties for the 1 and 2-dof body-pin and triangle-hyperpin systems defined above.

\begin{observation*}
\label{obs:bodypin_drp}
For the 1-dof body-pin graphs described above that are $(3,4)$-sparse, a $(3,4)$-tight canonical DR-plan exists where every node of a $(3,4)$-sparse graph satisfies one of the following: (1) its children are 2 proper vertex-maximal 1-dof graphs that intersect on another 1-dof graph; or (2) its children are all of the proper maximal 1-dof sub-graphs, pairwise sharing at most one body.
\end{observation*}

@@

As in Section \ref{sec:DRP},  a strong Church-Rosser property holds, making all canonical DR-plans optimal:
%
\begin{observation*}\label{rem:1dofcanon}
    When the input is independent, all $(3,4)$-tight canonical DR-plans are optimal. We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar-joint graphs discussed in Section \ref{sec:DRP}.
\end{observation*}

The above-mentioned algorithm exists because such $(3,4)$-tight graphs are matroidal and have a pebble game \cite{Lee:2007:PGA}.

@@

\begin{observation*}
\label{obs:2dof_case}
    For the 2-dof case, provided an analogous statement to Observation \ref{rem:1dofcanon} holds, then Theorem \ref{thm:1dofcase} holds for the 2-dof  systems.
\end{observation*}

@@

\begin{remark}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for the isostatic completion with minimum fan-in, (a reasonable measure of algebraic complexity), a more nuanced measure that treats solutions of 1-dof and 2-dof systems as 1 or 2 parameter families would no longer be optimized by the algorithm given in that proof. In particular, the complexity of the standard algorithm in the $k$-dof case would be exponential in $k$ (even if the case were matroidal and an optimal DR-plan is known).
\end{remark}

### 2015-05-16 doc/GMP2015/revision_2/s07_pinnedline.tex
Revision 2. 24 hours to submission.

The canonical DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve the structure of cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as \dfn{pinned line-incidence systems}.

\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property \cite{buehler2008nanomechanics}. \dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}. Both of these substances consist of a large number \dfn{microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with usually 3 other fibrils, where the \dfn{cross-linking} is like an incidence constraint that the crosslinked fibrils can slide against each other while remaining incident (see Figure~\ref{fig:crosslink}).

@@

In the case of microfibril cross-linking, each fibril is attached to some fixed larger organelle/membrane at one spot. Consequently, each fibril can be modeled as an edge of the graph, with the attachment being the corresponding pin. The two cross-linkings on the fibril are modeled as the two vertices in $V$ defining the edge.

@@

    The \dfn{canonical DR-plan} of $G$  is one in which the  children containing rigid subgraphs are connected isostatic vertex-maximal subgraphs of the parent.
\end{definition}

>>>>

The canonical DR-plan of Section \ref{sec:DRP} can additionally be applied to analyze and solve the structure of cross-linking collagen microfibrils in animals, cellulose microfibrils in plant cell walls, and actin filaments in the cytoskeleton by modeling these structures as a third type of qusecs, \dfn{pinned line-incidence systems}.

\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical properties \cite{buehler2008nanomechanics}. \dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}. Both of these substances consist of a large number \dfn{microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with usually 3 other fibrils, where the \dfn{cross-linking} is like an incidence constraint that the crosslinked fibrils can slide against each other while remaining incident (see Figure~\ref{fig:crosslink}).

@@

In the case of microfibril cross-linking, each fibril is attached to some fixed larger organelle/membrane at one site. Consequently, each fibril can be modeled as an edge of the graph, with the attachment being the corresponding pin. The two cross-linkings in which the fibril participates are modeled as the two vertices in $V$ defining the edge.

@@

    The \dfn{canonical DR-plan}  of $G$  is one in which the  child rigid subgraphs are connected, isostatic vertex-maximal subgraphs of the parent.
\end{definition}

### 2015-05-16 doc/GMP2015/revision_2/s08_conclusion.tex
Revision 2. 24 hours to submission.

We have clarified the main source of complexity for the optimal DR-plan and recombination problems. For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, which is a strong Church-Rosser property. This gives an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.

We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan. These results rely on a recent theory of Convex Cayley Configuration spaces. Relationships and reductions between these and previously studied problems were formally clarified.

@@

% \section{Open Problems}
\subsection{Open Problems}
\label{sec:appendix:b}
\label{sec:futurework}

One open problem comes from Section \ref{sec:DRP}:
\begin{openproblem}
    Is there a lower complexity algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
\end{openproblem}

A few natural  open questions concern the following common theme that runs through the paper:

\begin{openproblem}
    For fixed $k$ we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases under a modified measure for optimizing the DR-plan (Section~\ref{sec:table}). However, in the running time of all of these algorithms, $k$ appears in the exponent.
\end{openproblem}

>>>>

We have clarified the main source of complexity for the optimal DR-plan and recombination problems. For the former problem, when there are no overconstraints (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that any canonical DR-plan is guaranteed to be optimal, a strong Church-Rosser property. This gives an efficient (\candrpcomplexity) algorithm to find an optimal DR-plan that satisfies other desirable characteristics.

We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise indecomposable systems  at nodes of a DR-plan. These results rely on a recent theory of Convex Cayley Configuration spaces. Relationships and reductions between these and previously studied problems were formally clarified.

@@

% \section{Open Problems}
\subsection{Open Problems}
\label{sec:appendix:b}
\label{sec:futurework}
\label{sec:open}

The first set of problems are from Section \ref{sec:DRP}:
\begin{openproblem}
    Is there a more efficient algorithm than \candrpcomplexityv\ to find the canonical DR-plan of isostatic 2D bar-joint graphs?
\end{openproblem}

\begin{conjecture}
\label{conj:mfaisoptimal}
    The Modified Frontier Algorithm (MFA)~\cite{lomonosov2004graph} finds a canonical, and hence optimal, DR-plan.
\end{conjecture}

The difficulty of proving Conjecture~\ref{conj:mfaisoptimal} arises from the fact that MFA, although running in time $O(n^3)$, is a bottom-up algorithm, involving complex datastructures. However, a proof of optimality, even if it exists, would not be possible without the new notion of a canonical DR-plan at hand. The intuition for this conjecture comes from the similarity of the DR-plan generated by MFA to that of the sequential decomposition described in the proof of Theorem~\ref{theorem:algo_complexity}.  Since it is known \cite{lomonosov2004graph} that the DR-plan generated by MFA is cluster-minimal, an alternate conjecture is the following.

\begin{conjecture}
\label{conj:mfaisoptimal:rephrase}
For independent graphs, cluster-minimal DR-plans are optimal. In fact, for independent graphs, cluster-minimality and canonical are equivalent properties of a DR-plan.
\end{conjecture}

\begin{openproblem}
\label{open:sparsitymatroid}
Although generic rigidity is a property of graphs, and moreover,  in the case of qusecs, generic rigidity has a combinatorial sparsity and tightness-based characterization, the original definition of independence in the rigidity matroid requires an algebraic notion of independence of vectors of indeterminates over $\mathbf{R}$). Thus the definition of the DR-plan requires algebra over the reals.
In fact, the recursive decomposition problem is not tied to geometric constraint graphs or
an algebraic-geometric or mechanical notion of rigidity, and can be defined for any graph using
the notion of an abstract rigidity matroid \cite{graver93book}. This is  type of matroid with two additional matroid axioms,  abstract rigidity matroids can be defined in a purely graph-theoretic manner,  with no need for algebra in their definition. However, such abstract rigidity need not have a sparsity characterization. On the other hand, there are sparsity matroids that do not correspond to any notion of abstract rigidity.
However, when an abstract rigidity matroid is also a sparsity matroid, then the techniques of this
paper  directly apply and we can obtain purely combinatorially defined recursive decompositions of graphs.
\end{openproblem}

A few natural  open questions concern the following common theme that runs through the optimal recombination and later sections of the paper:

\begin{openproblem}
    For fixed $k$, we have polynomial time optimal DR-planning (Section~\ref{sec:DRP}), recombination (modification) in the presence of $k$ overconstraints, optimal modification for decomposition {\sl OMD}$_k(G)$, when at most $k$ constraints are removed (Section~\ref{sec:recomb}) and also optimal completion using at most $k\le 2$ constraints in the body-pin and triangle-multipin cases for a somewhat different optimization of the DR-plan (Section~\ref{sec:table}). However, in the running time of all of these algorithms, $k$ appears in the exponent.  Can $k$ be removed from the exponent?
\end{openproblem}
